---
title: "Assignment 2"
author: "KR"
date: "19 feb 2019"
output: html_document
---

```{r, wd, libraries}
#setwd
setwd("C:/Users/Reka/Desktop/CS 4/R")
#libraries
library(tidyverse)
library(rethinking)

```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):

- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

###1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?

- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results

```{r, grid approximation}

#define grid
p_grid <- seq(from= 0, to = 1, length.out = 20)

#define prior
prior <- rep(1,6)

#compute likelihood at each value in grid
likelihood <- dbinom(3 , size = 6, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood * prior

#standardise the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

#display distribution
plot(p_grid, posterior, type ="b", xlab ="probability of correct answer", ylab = "posterior probability")


```

- Then implement a quadratic approximation (hint check paragraph 2.4.2!).

```{r, quadratic approximation}

globe.qa <- map( 
  alist( 
    k ~ dbinom(6,p) ,# binomial likelihood  
    p ~ dunif(0,1) # uniform prior
    ), data=list(k=3) ) 

#display summary of quadratic approximation
precis(globe.qa)
```

- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r, more than chance (0.5)}
sum ( posterior[p_grid > 0.5]) #There is 0.5 probability that Riccardo knows more than chance.

```

The distribution is centered around chance (0.5) - most likely (estimated knowledge), but uncertain - large interval. The quadratic appr. shows the same, assuming the posterior is gaussian - maximized at 0.5 with a std of 0.2; 

###2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.

2a. Produce plots of the prior, and posterior for each teacher.

```{r}
#function to grid approx. 
likelihood_p <- function(p, count, size) {
  return(dbinom(prob = p, x = count, size = size))
} # likelihood of p given the data

p_grid <- seq(0,1, length.out = 200)
prior_uni <- rep(1,200)

right_RIC <- 3 
right_KRI <- 2
right_JOS <- 160
right_MIK <- 66

total_RIC <- 6
total_KRI <- 2
total_JOS <- 198
total_MIK <- 132

```

```{r}
#Riccardo
RIC <- likelihood_p(p_grid, 
                      count = right_RIC,
                      size = total_RIC) * prior_uni

RIC_std <- RIC/sum(RIC)

#Kristian
KRI <- likelihood_p(p_grid, 
                      count = right_KRI,
                      size = total_KRI) * prior_uni

KRI_std <- KRI/sum(KRI)

#Josh
JOS <- likelihood_p(p_grid, 
                      count = right_JOS,
                      size = total_JOS) * prior_uni

JOS_std <- JOS/sum(JOS)

#Mikkel
MIK <- likelihood_p(p_grid, 
                      count = right_MIK,
                      size = total_MIK) * prior_uni

MIK_std <- MIK/sum(MIK)

```

```{r, more than chance}
sum ( RIC_std[p_grid > 0.5])
sum ( KRI_std[p_grid > 0.5])
sum ( JOS_std[p_grid > 0.5])
sum ( MIK_std[p_grid > 0.5])

p_grid[ which.max(RIC_std) ]
p_grid[ which.max(KRI_std) ]
p_grid[ which.max(MIK_std) ]
p_grid[ which.max(JOS_std) ]

```

```{r, plots}
#RIC
plot(p_grid, RIC_std, type = "l", xlab ="probability of correct answer (Riccardo)", ylab = "posterior probability")
#KRI
plot(p_grid, KRI_std, type = "l",xlab ="probability of correct answer (Kristian)", ylab = "posterior probability")
#JOS
plot(p_grid, JOS_std, type = "l", xlab ="probability of correct answer (Josh)", ylab = "posterior probability")
#MIK
plot(p_grid, MIK_std, type = "l", xlab ="probability of correct answer (Mikkel)", ylab = "posterior probability")
```

According to the plot, Kristian seems to be the best of the teachers.
```{r, single plot}
plot <- ggplot() +
  aes(x = p_grid, y = MIK_std) +
  geom_line(aes(col = 'red'), size = 1) + ylab('posterior probability') + xlab('Probability of correct answer')
plot <- plot + geom_line(aes(y= RIC_std), colour="green", size = 1)
plot <- plot + geom_line(aes(y= KRI_std), colour="orange", size = 1)
plot <- plot + geom_line(aes(y= JOS_std), colour="blue", size = 1) + theme_classic()
plot
```

Both (Mikkel + Riccardo) distributions are centered around 0.5, saying that they are likely to answer half of the questions correctly. But Mikkel's model is much more certain in this case, that the score will end up very close to the center giving no chance to either end (perfect or null scores). 


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r, Prior with normal distribution (0.8, 0.2)}

prior_p <- dnorm(p_grid, 0.8, 0.2)
#plot(p_grid, prior_p , type = "l", xlab ="probability of correct answer (prior) ", ylab = "posterior probability")

#Riccardo
RIC <- likelihood_p(p_grid, 
                      count = right_RIC,
                      size = total_RIC) * prior_p

RIC_std_2 <- RIC/sum(RIC)

#Kristian
KRI <- likelihood_p(p_grid, 
                      count = right_KRI,
                      size = total_KRI) * prior_p

KRI_std_2 <- KRI/sum(KRI)

#Josh
JOS <- likelihood_p(p_grid, 
                      count = right_JOS,
                      size = total_JOS) * prior_p

JOS_std_2 <- JOS/sum(JOS)

#Mikkel
MIK <- likelihood_p(p_grid, 
                      count = right_MIK,
                      size = total_MIK) * prior_p

MIK_std_2 <- MIK/sum(MIK)

#Find peaks
p_grid[ which.max(RIC_std_2) ]
p_grid[ which.max(KRI_std_2) ]
p_grid[ which.max(MIK_std_2) ]
p_grid[ which.max(JOS_std_2) ]


```

```{r, plots}
plot(p_grid, RIC_std_2, type = "l", xlab ="probability of correct answer (Riccardo)", ylab = "posterior probability")
plot(p_grid, KRI_std_2, type = "l",xlab ="probability of correct answer (Kristian)", ylab = "posterior probability")
plot(p_grid, JOS_std_2, type = "l", xlab ="probability of correct answer (Josh)", ylab = "posterior probability")
plot(p_grid, MIK_std_2, type = "l", xlab ="probability of correct answer (Mikkel)", ylab = "posterior probability")

```

```{r, single plot}
#For fun
plot <- ggplot() +
  aes(x = p_grid, y = MIK_std_2 ) +
  geom_line(aes(col = 'red'))
plot <- plot + geom_line(aes(y= RIC_std_2), colour="green")
plot <- plot + geom_line(aes(y= KRI_std_2), colour="orange")
plot <- plot + geom_line(aes(y= JOS_std_2), colour="blue")
plot
```

The prior is no longer flat but a distribution with the mean/peak of 0.8 and a std of 0.2.Kinda similar, all of the teacher's curves are pushed towards the mean of the prior. Still assumes Kristian to be the best, but no longer with a perfect score (1). Josh's curve is not really affected by the new prior. Riccardo
Riccardo has higher probability to have above average scores now, but still shows great uncertainty compared to others. Compared to Mikkel, Riccardo now has higher probability of having more correct answers.

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r, uni+*100}

add_right_RIC <- 3 *100
add_right_KRI <- 2 *100
add_right_JOS <- 160 *100
add_right_MIK <- 66 *100
 
add_total_RIC <- 6 *100
add_total_KRI <- 2 *100
add_total_JOS <- 198 *100
add_total_MIK <- 132 *100

#Riccardo
add_RIC <- likelihood_p(p_grid, 
                      count = add_right_RIC,
                      size = add_total_RIC) * prior_uni

add_RIC_std <- add_RIC/sum(add_RIC)

#Kristian
add_KRI <- likelihood_p(p_grid, 
                      count = add_right_KRI,
                      size = add_total_KRI) * prior_uni

add_KRI_std <- add_KRI/sum(add_KRI)

#Josh
add_JOS <- likelihood_p(p_grid, 
                      count = add_right_JOS,
                      size = add_total_JOS) * prior_uni

add_JOS_std <- add_JOS/sum(add_JOS)

#Mikkel
add_MIK <- likelihood_p(p_grid, 
                      count = add_right_MIK,
                      size = add_total_MIK) * prior_uni

add_MIK_std <- add_MIK/sum(add_MIK)

#Max point
p_grid[ which.max(add_RIC_std) ]
p_grid[ which.max(add_KRI_std) ]
p_grid[ which.max(add_MIK_std) ]
p_grid[ which.max(add_JOS_std) ]

```

```{r, plots_uni+*100}
#Plots
plot(p_grid, add_RIC_std, type = "l", xlab ="probability of correct answer (Riccardo)", ylab = "posterior probability")
plot(p_grid, add_KRI_std, type = "l",xlab ="probability of correct answer (Kristian)", ylab = "posterior probability")
plot(p_grid, add_JOS_std, type = "l", xlab ="probability of correct answer (Josh)", ylab = "posterior probability")
plot(p_grid, add_MIK_std, type = "l", xlab ="probability of correct answer (Mikkel)", ylab = "posterior probability")
```

```{r, single plot_uni+*100}
plot <- ggplot() +
  aes(x = p_grid, y = add_MIK_std ) +
  geom_line(aes(col = 'red'))
plot <- plot + geom_line(aes(y= add_RIC_std), colour="green")
plot <- plot + geom_line(aes(y= add_KRI_std), colour="orange")
plot <- plot + geom_line(aes(y= add_JOS_std), colour="blue")
plot
```

```{r, (0.8, 0.2) }

prior_skep<- dnorm(p_grid, 0.8, 0.02)

#Riccardo
add_RIC <- likelihood_p(p_grid, 
                      count = add_right_RIC,
                      size = add_total_RIC) * prior_p
add_RIC_std_2 <- add_RIC/sum(add_RIC)


#KRI
add_KRI <- likelihood_p(p_grid, 
                      count = add_right_KRI,
                      size = add_total_KRI) * prior_p

add_KRI_std_2 <- add_KRI/sum(add_KRI)


#JOS
add_JOS <- likelihood_p(p_grid, 
                      count = add_right_JOS,
                      size = add_total_JOS) * prior_p

add_JOS_std_2 <- add_JOS/sum(add_JOS)


#MIK
add_MIK <- likelihood_p(p_grid, 
                      count = add_right_MIK,
                      size = add_total_MIK) * prior_p

add_MIK_std_2 <- add_MIK/sum(add_MIK)


#Max point
p_grid[ which.max(add_RIC_std_2) ]
p_grid[ which.max(add_KRI_std_2) ]
p_grid[ which.max(add_MIK_std_2) ]
p_grid[ which.max(add_JOS_std_2) ]


```

```{r, plots - (0.8, 0.2) + *100}
plot(p_grid, add_RIC_std_2, type = "l", xlab ="probability of correct answer (Riccardo)", ylab = "posterior probability")
plot(p_grid, add_KRI_std_2, type = "l", xlab ="probability of correct answer (Kristian)", ylab = "posterior probability")
plot(p_grid, add_JOS_std_2, type = "l", xlab ="probability of correct answer (Josh)", ylab = "posterior probability")
plot(p_grid, add_MIK_std_2, type = "l", xlab ="probability of correct answer (Mikkel)", ylab = "posterior probability")
```

```{r, single plot - (0.8, 0.2) * 100}

plot <- ggplot() +
  aes(x = p_grid, y = add_MIK_std_2 ) +
  geom_line(aes(col = 'red'))
plot <- plot + geom_line(aes(y= add_RIC_std_2), colour="green")
plot <- plot + geom_line(aes(y= add_KRI_std_2), colour="orange")
plot <- plot + geom_line(aes(y= add_JOS_std_2), colour="blue")
plot

```

when the prior is uniform and we have such a large number of observations the posterior is only one value, we are almost certain of the teacher's performance due to the large number of tests and the uniformity of priors. The uniform priors aren't capturing any of the "large-world" uncertainty 
when the prior is normal and we have such a large number of observations the posterior is only one value, we are almost certain of the teacher's performance due to the large number of tests, as the evidence increases the model relies less on the priors.


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

Have 'pessimistic' priors, meaning that if we'd ask the teachers cogsci questions it's highly unlikely (10% chance) that the rate of their correct answers would be bigger than chance.
//set mean to 0.5 (there are only yes-no answers) with low std.

```{r, echo= TRUE}
prior3 <- dnorm(1:10, 0.5 , 0.1)
```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.

```{r}
library(pacman)
p_load(DT)
p_load(brms)

d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1, data = subset(d,Teacher=="RF"), prior = prior("uniform(0,1)", class = "Intercept"),family=binomial)
plot(FlatModel)

PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"),family=binomial)
plot(PositiveModel)

SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"),family=binomial)
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?

## Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. 
N.B. You can choose which prior to use for the analysis of last year's data.


Questions to be answered (but see guidance below):
###1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models

In freq. models given the number and nature of observations the mean of the given data would be the prediction the model would make for future data. The error ends up being a point estimate, not accounting for much of the uncertainty. 
In bayesian models we can account for the full probabilty distribution related to each probable value of a variable, while also taking into account the prior knowledge we have and the likelihood of the scores we got, thus describing the full range of possible estimates. We can make more detailed predictions, for each probable instance of the variable examined.


###2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions


Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)

```{r, way 2}
p_grid <- seq(0,1, length.out = 200)

#Riccardo
RIC_samples = sample(p_grid, prob = RIC_std_2, size = 1e4, replace = TRUE) # sample from the posterior
dens(RIC_samples) 
#Posterior Predictive Distribution
RIC_pre <- rbinom( 1e4 , size=10 , prob= RIC_samples)

#Kristian
KRI_samples = sample(p_grid, prob = KRI_std_2, size = 1e4, replace = TRUE) # sample from the posterior
dens(KRI_samples) 
#PPD
KRI_pre <- rbinom( 1e4 , size=12 , prob= KRI_samples)


#Josh
JOS_samples = sample(p_grid, prob = JOS_std_2, size = 1e4, replace = TRUE) # sample from the posterior
dens(JOS_samples) 
#PPD
JOS_pre <- rbinom( 1e4 , size=172 , prob= JOS_samples)
set.seed(70)


#Mikkel
MIK_samples = sample(p_grid, prob = MIK_std_2, size = 1e4, replace = TRUE) # sample from the posterior
dens(MIK_samples)
#PPD
MIK_pre <- rbinom( 1e4 , size=65 , prob= MIK_samples)

```

```{r, histograms}

simplehist(RIC_pre, xlab = "Posterior Prediction (Riccardo)")
        abline(v = 9, col = "red")
        
simplehist(KRI_pre, xlab = "Posterior Prediction (Kristian)")
        abline(v = 8, col = "red")
           
simplehist(JOS_pre, xlab = "Posterior Prediction (Josh)")
        abline(v = 148, col = "red")

simplehist(MIK_pre, xlab = "Posterior Prediction (Mikkel)")
        abline(v = 34, col = "red")

```


### Depending on time: Questions from the handbook
2H1. Suppose there are two species of panda bear. Both are equally common in the wild and live in the same places. They look exactly alike and eat the same food, and there is yet no genetic assay capable of telling them apart. They differ however in their family sizes. Species A gives birth to twins 10% of the time, otherwise birthing a single infant. Species B births twins 20% of the time, otherwise birthing singleton infants. Assume these numbers are known with certainty, from many years of field research.
Now suppose you are managing a captive panda breeding program. You have a new female panda of unknown species, and she has just given birth to twins. What is the probability that her next birth will also be twins?

2H2. Recall all the facts from the problem above. Now compute the probability that the panda we have is from species A, assuming we have observed only the first birth and that it was twins.

2H3. Continuing on from the previous problem, suppose the same panda mother has a second birth and that it is not twins, but a singleton infant. Compute the posterior probability that this panda is species A.

2H4. A common boast of Bayesian statisticians is that Bayesian inference makes it easy to use all of the data, even if the data are of different types. So suppose now that a veterinarian comes along who has a new genetic test that she claims can identify the species of our mother panda. But the test, like all tests, is imperfect. This is the information you have about the test:
- The probability it correctly identifies a species A panda is 0.8.
- The probability it correctly identifies a species B panda is 0.65.
The vet administers the test to your panda and tells you that the test is positive for species A. First ignore your previous information from the births and compute the posterior probability that your panda is species A. Then redo your calculation, now using the birth data as well.
