---
title: "Untitled"
author: "Réka Keresztényi"
date: "August 14, 2020"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#workingdirect.
setwd("C:/Users/Reka/Desktop/m")
#libraries
pacman::p_load(tidyverse,ggplot2,lmerTest,pastecs,dplyr)
library(brms)
library(readr)
library(plyr)
library(ggplot2)
library(rethinking)
library(tidyverse)
library(rstanarm) 
library(tidybayes) 
library(modelr) 
library(gganimate) 
library(reshape)
library(psych)

library("rstan")
library("devtools")
#devtools::install_github("stan-dev/cmdstanr")
#install.packages(c("coda","mvtnorm","devtools","loo","dagitty"))
#devtools::install_github("rmcelreath/rethinking")

```

```{r}

#data
data <- read_csv("LaughData.csv")
describe(data)
 
#data2 <- na.omit(data) #no NA
#subset + have average Hitrate in both conditons (strangers + friends)
data2 <- aggregate(data[, 7], list(data$Subject, data$Country, data$Condition, data$Sex), mean)
data3 <- aggregate(data[, 7], list(data$Country, data$Condition), mean)

colnames(data2) <- c("Subject","Country","Condition","Gender", "Hit")
colnames(data3) <- c("Country","Condition","Hit")

mydata <- data3
ggplot(mydata,aes(Hit))+geom_density() #look for prior to explain it
describe(data3)

plot( Hit ~ Condition , data=data3) + geom_boxplot() #hit rate higher in friends
ggplot(data3, aes(Condition, Hit,group=Condition)) + geom_boxplot()

#plots
ggplot(data3, aes(x=Condition, y=Hit, group=Condition)) + 
  geom_bar(stat='identity') + facet_grid(~Country) + facet_wrap(~ Country)

ggplot(data2, aes(x=Condition, y=Hit, group=Condition)) + 
  geom_bar(stat='identity')  + facet_grid(~Country) + facet_wrap(~ Country)

#models
glmer(Hit ~ Condition*Country + (1 + 1Subject))

```

```{r, include=FALSE}

#dependent + independent variables
mydata$x1 = as.factor(mydata$Condition) 
mydata$x2 = mydata$Country 
mydata$y = mydata$Hit

x_lab = "Condition" 
y_lab = "Hitrate"

```

```{r, include=FALSE}

a_prior = 0.7   
a_prior_max = 1 
# get the value of 1 sd by finding the maximum amount our data can vary from the mean (a_prior_max - a_prior) and divide that in half.
a_sd = (a_prior_max - a_prior) / 2  

#we don't have a strong hypothesi so we set the mean of the effect size parameters to be 0. In the absence of other information, we set the sd to be the same as for the control condition.
b1_prior = 0  
b1_sd = a_sd   

```

```{r, include=FALSE}

#generate the prior distribution
m_prior = stan_glm(y ~ x1*x2, data = mydata,
  prior_intercept = normal(a_prior, a_sd, autoscale = FALSE),
  prior = normal(b1_prior, b1_sd, autoscale = FALSE),
  prior_PD = TRUE
)
#check
pp_check(m_prior)
precis(m_prior)

# Create the dataframe with fitted draws
prior_draws = mydata %>% 
                data_grid(x1, x2) %>% 
                add_fitted_draws(m_prior, n = 5, seed = 12345) 

#function_plotting
static_prior_plot_4 = function(prior_draws) {
  ggplot(prior_draws, aes(x = x1, y = .value, col = x2, fill = x2, group = x2)) +
    geom_bar(stat='identity', position='dodge') +
    facet_grid(cols = vars(.draw)) +
    #facet_wrap(~x2)+
    theme(strip.background = element_blank(),
          strip.text.y = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5)) +
    labs(x=x_lab, y=y_lab) + # axes labels
    ggtitle("Five sample draws from the priors")
}

static_prior_plot_4(prior_draws)

 
```

Hops visualization:

```{r, include=FALSE}

n_draws = 50 #trialsnr.
frames_per_second = 2.5 
# 2.5 frames per second (400ms) is the recommended speed for the HOPs visualization.

# Add more prior draws to the data frame for the visualization
more_prior_draws <- rbind(
                mydata %>%
                data_grid(x1,x2) %>%
                add_fitted_draws(m_prior, n = n_draws - 5, seed = 12345))

#function
HOPs_plot_4 = function(data) {
  ggplot(data, aes(x = x1, y = .value, col = x2, fill = x2, group = x2)) + #do not change
    geom_bar(stat='identity', position='dodge') + #do not change from stat='identity'. Fill and line aesthetics may be modified here, see ggplot2 documentation
    transition_states(.draw, transition_length = 1, state_length = 1) + # gganimate code to animate the plots. Do not change
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #rotates the x-axis text for readability
    coord_cartesian(ylim = c(min(mydata$y, na.rm=T), max(mydata$y, na.rm=T))) + # sets axis limits
    # scale_x_discrete(limits=c("before","after")) + #manually set the order of the x-axis levels
    labs(x=x_lab, y=y_lab) # axes labels
}

# this function is defined in 'plotting_functions.R', if you wish to customize the aesthetics.
prior_HOPs = animate(HOPs_plot_4(more_prior_draws), nframes = n_draws * 2, fps = frames_per_second)
prior_HOPs
```

```{r, include=FALSE}
#posterior
m = stan_glm(y ~ x1*x2, data = mydata,
  prior_intercept = normal(a_prior, a_sd, autoscale = FALSE),
  prior = normal(b1_prior, b1_sd, autoscale = FALSE)
)

pp_check(m)
precis(m)
summary(m, digits=3)



```


Here is a summary of the model fit - Rhat values are very close to 1.0 - implying the model is a good fit for our data. 

For each x position in the fit grid, we can get the posterior mean estimates and 95% credible intervals from the model:

```{r, include=FALSE}
# Create the dataframe with fitted draws
fit = mydata %>%#
      data_grid(x1,x2) %>% 
      add_fitted_draws(m) %>% 
      mean_qi(.width = .95) #add 95% credible intervals
#function
static_post_plot_4 = function(fit) {
  ggplot(fit, aes(x = x1, y = .value, fill = x2, group = x2)) +
    geom_bar(stat='identity', position='dodge') +
    geom_errorbar(aes(ymin = .lower, ymax = .upper), position = position_dodge(width = .9), width = .2) + 
    coord_cartesian(ylim = c(min(mydata$y, na.rm=T), max(mydata$y, na.rm=T))) +
    labs(x=x_lab, y=y_lab)
}

# Plot the posterior draws
static_post_plot_4(fit)

```


Sampling from the posterior
To get a better visualization of the uncertainty remaining in the posterior results, we can use animated HOPs for this graph as well. (replace `m_prior` with `m`)

```{r, include=FALSE}
p = mydata %>% 
  data_grid(x1, x2) %>% 
  add_fitted_draws(m, n = n_draws, seed = 12345) 
post_HOPs = animate(HOPs_plot_4(p), nframes = n_draws * 2, fps = frames_per_second) 
post_HOPs
```

Comparing the prior and posterior
- adding information to the model adds more certainty to our estimates

```{r , include=FALSE}
prior_HOPs
post_HOPs
```


